---
title: "Hotel Bookings"
output: html_document
date: "2023-12-29"
---
# Analysis Goal: Explore the data to gain insights on running promotions/deals.

## Setup Packages
```{r install packages}
#install.packages("tidyverse")
#install.packages("skimr")
#install.packages("janitor")
#install.packages("ggplot2)
```

```{r load packages}
library(tidyverse)
library(skimr)
library(janitor)
library(ggplot2)
```

## Import Data
```{r load dataset}
hotel_bookings <- read_csv("hotel_bookings.csv")
```

## Explore Data
```{r}
head(hotel_bookings)
```

```{r}
str(hotel_bookings)
```

```{r}
colnames(hotel_bookings)
```

## How early should you run promotions for city hotel rooms vs resort hotel rooms

### Types of hotel rooms in dataset
```{r}
hotel_bookings %>% 
  group_by(hotel) %>% 
  summarize(count = n()) %>% 
  print()
```

### Find average lead time for city vs resort
```{r}
leadtime_summary <- hotel_bookings %>% 
  group_by(hotel) %>% 
  summarize(
    Avg_lead_time = mean(lead_time),
    Max_lead_time = max(lead_time),
    Min_lead_time = min(lead_time)
  )
head(leadtime_summary)
```
### Visual with count of lead_time for specified bins.
```{r}
leadtime_per_hotel <- hotel_bookings %>% 
  group_by(hotel) %>% 
  summarize(
    count_0_100 = sum(lead_time >= 0 & lead_time <= 100),
    count_101_200 = sum(lead_time > 100 & lead_time <= 200),
    count_201_300 = sum(lead_time > 200 & lead_time <= 300),
    count_301_400 = sum(lead_time > 300 & lead_time <= 400),
    count_401_500 = sum(lead_time > 400 & lead_time <= 500),
    count_501_600 = sum(lead_time > 500 & lead_time <= 600),
    count_601_700 = sum(lead_time > 600 & lead_time <= 700)
  )

# Reshape data for ggplot
leadtime_per_hote_reshaped <- tidyr::gather(leadtime_per_hotel, key = "lead_time_range", value = "count", -hotel)

# Create a bar plot
ggplot(leadtime_per_hote_reshaped, aes(x = hotel, y = count, fill = lead_time_range)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Lead Time Distribution",
       x = "Hotel",
       y = "# Bookings") +
  scale_fill_brewer(palette = "Set3")
```
### Takeaway
- Avg lead time for City Hotel: 110 days
- Avg lead time for Resort Hotel: 93 days
- Double bookings for City vs Resort Hotels 

## Number of children vs lead time
```{r}
ggplot(data = hotel_bookings) +
  geom_point(mapping = aes(x = lead_time, y = children))+
  labs(title = "Number of Children by Lead Time",
       x = "Lead Time",
       y = "# Children")
```
### Takeaway
- Bookings with 0 children have longer booking lead time vs bookings with more than 1 children. 

```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = hotel, fill = hotel)) +
  facet_wrap(~children) +
  labs(title = "Number of Children by Hotel",
       x = "Hotel",
       y = "# Children") +
  scale_fill_brewer(palette = "Set3")
```
## Weekend vs Weekday bookings

```{r}
hotel_bookings %>% 
  summarize(
    total_nights = sum(stays_in_weekend_nights + stays_in_week_nights), 
    total_weekend_nights = sum(stays_in_weekend_nights),
    total_week_nights = sum(stays_in_week_nights),
    avg_night_per_booking = mean(stays_in_weekend_nights + stays_in_week_nights),
    avg_weekend_nights_per_booking = mean(stays_in_weekend_nights),
    avg_week_nights_per_booking = mean(stays_in_week_nights),
  )
```
### Total Weekend vs Total Week Nights
```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = factor(1), y = stays_in_weekend_nights, fill = "Weekend Nights"), stat = "identity") +
  geom_bar(mapping = aes(x = factor(2), y = stays_in_week_nights, fill = "Week Nights"), stat = "identity") +
  labs(title = "Total Weekend vs Total Week Nights",
       x = NULL,  # Remove x-axis label
       y = "Count") +
  scale_x_discrete(labels = c("Weekend Nights", "Week Nights")) +
  scale_fill_brewer(palette = "Set3")

```
### Average Weekend Nights vs Avg Week Nights per Booking
```{r}
avg_nights_comparison <- hotel_bookings %>%
  summarise(
    avg_weekend_nights = mean(stays_in_weekend_nights),
    avg_week_nights = mean(stays_in_week_nights)
  )

ggplot(data = avg_nights_comparison) +
  geom_bar(mapping = aes(x = factor(1), y = avg_weekend_nights, fill = "Avg Weekend Nights"), stat = "identity") +
  geom_bar(mapping = aes(x = factor(2), y = avg_week_nights, fill = "Avg Week Nights"), stat = "identity") +
  labs(title = "Average Weekend Nights vs Avg Week Nights per Booking",
       x = NULL,  # Remove x-axis label
       y = "Average Nights") +
  scale_x_discrete(labels = c("Avg Weekend Nights", "Avg Week Nights")) +
  scale_fill_brewer(palette = "Set3")

```
## Number of Transactions for Various Distribution Types
```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel, fill = distribution_channel)) +
  labs(title = "Num Transactions per Distribution Type",
       x = "Distribution Channels",
       y = "Count") +
  scale_fill_brewer(palette = "Set3")
```


```{r}
ggplot(data = hotel_bookings) +
  geom_bar(mapping = aes(x = distribution_channel, fill = market_segment)) +
  labs(title = "Num Transactions per Distribution Type split based on Market Segment",
       x = "Distribution Channels",
       y = "Count") +
  scale_fill_brewer(palette = "Set3")
```





